Simulation study in rectangular domain

### 2024-04-30 R/simulation_study/simulation_study_fjords.Rmd
add calculation of biais and rmse in simulation_study_rect.Rmd

```{r}
#number of batch
N_BATCH=10

@@

#degree of freedom in tensor splines
SP_DF=5
```

@@

@@

```{r}
res=sim_theta_CRCVM(ftau=ftau_bump,fomega=fomega,fnu=fnu_constant,v0=v0,x0=x0[2,],times=seq(0,5*24,by=1/60),land=land,verbose=FALSE)

@@

data_fjords_hf_ne <- foreach(i=1:N_BATCH,.packages=c("progress","MASS","sf","foreach")) %dopar% {
  data=foreach (i=1:N_ID,.combine='rbind') %do% {

@@

```{r percentage of low error samples that reached land}
count=0
for (k in 1:length(names(data_rect_hf_ne))) {

@@

plot_samples_fjords_hf_ne=ggplot()+geom_sf(data=land$geometry,fill=NA)+coord_sf(datum=st_crs(32626))+
  geom_point(data=data_fjords_hf_ne$BATCH1,mapping=aes(y1,y2,color=ID),size=0.1)+
  geom_path(data=data_fjords_hf_ne$BATCH1,mapping=aes(y1,y2,color=ID),size=0.1)+
  geom_point(data = data_fjords_hf_ne$BATCH2%>% filter(!duplicated(ID)),
             aes(x = y1, y = y2), shape = 3, size = 4, col = "red")+
  xlab("x") + ylab("y")

@@

crcvm_fjords_hf_ne1<- foreach(i=1:N_BATCH,.packages=c("smoothSDE")) %dopar% {
  #estimation with tensor splines
  formulas <- list(mu1=~1,mu2=~1,tau=~s(ID,bs="re"),
                 nu=~s(ID,bs="re"),omega=~te(theta,ExpShore,k=SP_DF,bs="cs"))
  par0 <- c(0,0,1,1,0)
  m <- SDE$new(formulas = formulas,data = data_fjords_hf_ne[[paste("BATCH",k,sep="")]],type = "RACVM",
                      response = c("y1","y2"),par0 = par0,fixpar=c("mu1","mu2"),
                      other_data=list("H"=H_hf))
  m$fit(method="BFGS")
  m 
}

>>>>

```{r}
#number of batch
N_BATCH=1

@@

#degree of freedom in tensor splines
SP_DF=3
```

@@

```{r}
sp_coeff_Dshore
```

```{r}
sp_coeff_ExpShore
```

@@

```{r illustrative sample}
res=sim_theta_CRCVM(ftau=ftau_constant,fomega=fomega,fnu=fnu_constant,v0=v0,x0=x0[2,],times=seq(0,5*24,by=1/60),land=land,verbose=FALSE)

@@

data_fjords_hf_ne <- foreach(k=1:N_BATCH,.packages=c("progress","MASS","sf","foreach")) %dopar% {
  data=foreach (i=1:N_ID,.combine='rbind') %do% {

@@

```{r percentage of low error samples that reached land}
count=0
for (k in 1:length(names(data_fjords_hf_ne))) {

@@

plot_samples_fjords_hf_ne=ggplot()+geom_sf(data=land$geometry,fill=NA)+coord_sf(datum=st_crs(32626))+
  geom_point(data=data_fjords_hf_ne$BATCH1,mapping=aes(y1,y2,color=ID),size=0.1)+
  geom_path(data=data_fjords_hf_ne$BATCH1,mapping=aes(y1,y2,color=ID),size=0.1)+
  geom_point(data = data_fjords_hf_ne$BATCH1%>% filter(!duplicated(ID)),
             aes(x = y1, y = y2), shape = 3, size = 4, col = "red")+
  xlab("x") + ylab("y")

@@

crcvm_fjords_hf_ne1<- foreach(k=1:N_BATCH,.packages=c("smoothSDE")) %dopar% {
  #estimation with tensor splines
  formulas <- list(mu1=~1,mu2=~1,tau=~s(ID,bs="re"),
                 nu=~s(ID,bs="re"),omega=~te(theta,ExpShore,k=SP_DF,bs="cs"))
  par0 <- c(0,0,1,1,0)
  m <- SDE$new(formulas = formulas,data = data_fjords_hf_ne[[paste("BATCH",k,sep="")]],type = "RACVM",
                      response = c("y1","y2"),par0 = par0,fixpar=c("mu1","mu2"),
                      other_data=list("H"=H_hf))
  m$fit(method="BFGS")
  m 
}

### 2024-04-30 R/simulation_study/simulation_study_rect.Rmd
add calculation of biais and rmse in simulation_study_rect.Rmd

```{r global variables}
#number of batch
N_BATCH=100

@@

```{r measurement error}
#no measurement error (we cannot put 0 because only the log of this quantity is used)
sigma_obs=0.01
H_hf=array(rep(sigma_obs^2*diag(2),n_hf*N_ID),dim=c(2,2,n_hf*N_ID))

@@

plot_samples_rect_hf_ne=ggplot()+geom_sf(data=domain$geometry,fill="grey")+coord_sf(datum=st_crs(32626))+
  geom_point(data=data_rect_hf_ne$BATCH1,mapping=aes(y1,y2,color=ID,shape=ID),size=0.1)+
  geom_path(data=data_rect_hf_ne$BATCH1,mapping=aes(y1,y2,color=ID,shape=ID),size=0.1)+
  geom_point(data = data_rect_hf_ne$BATCH1%>% filter(!duplicated(ID)),
             aes(x = y1, y = y2), shape = 3, size = 4, col = "red")+
  xlab("x") + ylab("y")

ggsave(filename="samples_rect_hf_ne_omega.png",plot=plot_samples_rect_hf_ne,width=10,height=5)
ggplotly(plot_samples_rect_hf_ne)
```

@@

#### Violin plot of the estimated spline coefficients for one batch

@@

for (k in 1:N_BATCH) {
  m<-crcvm_rect_hf_ne1[[paste("BATCH",k,sep="")]]
  coeffs=m$coeff_re()
  coeff_names=rownames(coeffs)
  coeff_values=as.numeric(coeffs)
  sub_index=grepl("omega",coeff_names)
  sub_coeffs_df=data.frame("estimate"=coeff_values[sub_index],"coeff_name"=factor(coeff_names[sub_index]))
  coeffs_df=rbind(coeffs_df,sub_coeffs_df)
}

@@

@@

```{r plot estimates te DistanceShore for one batch of low error samples}
xmin=list("theta"=-pi,"DistanceShore"=0.3)
xmax=list("theta"=pi,"DistanceShore"=2)
res=crcvm_rect_hf_ne2$BATCH1$get_all_plots(baseline=NULL,model_name="crcvm_rect_hf_ne2",xmin=xmin,xmax=xmax,
                                    xlabel=xlabel,link=link,show_CI="pointwise",true_smooths=true_smooths,save=TRUE)
res$fe_omega_theta_DistanceShore
```

@@

for (k in 1:N_BATCH) {
  m<-crcvm_rect_hf_ne2[[paste("BATCH",k,sep="")]]
  coeffs=m$coeff_re()
  coeff_names=rownames(coeffs)
  coeff_values=as.numeric(coeffs)
  sub_index=grepl("omega",coeff_names)
  sub_coeffs_df=data.frame("estimate"=coeff_values[sub_index],"coeff_name"=factor(coeff_names[sub_index]))
  coeffs_df=rbind(coeffs_df,sub_coeffs_df)
}

@@

p=ggplotly(p)
saveWidget(p, file ="crcvm_rect_hf_ne2/all_batches_coeffs_crcvm_rect_hf_ne2.html")

p
```

@@

@@

@@

atw=list("ExpShore"=fExpShore,"AngleNormal"=fangle)
sim_data=crcvm_rect_hf_ne2$BATCH1$simulate(z0=x0[1,],data=crcvm_rect_hf_ne2$BATCH1$data(),atw=atw,land=domain)

>>>>

```{r global variables}
#number of batch
N_BATCH=10

@@

```{r measurement error}
#no measurement error (we cannot put 0 because only the log of this quantity is used)
sigma_obs=0.005
H_hf=array(rep(sigma_obs^2*diag(2),n_hf*N_ID),dim=c(2,2,n_hf*N_ID))

@@

plot_samples_rect_hf_ne=ggplot()+geom_sf(data=domain$geometry,fill="grey")+coord_sf(datum=st_crs(32626))+
  geom_point(data=data_rect_hf_ne$BATCH1,aes(p1,p2),shape=5,size=0.1,col="red")+
  geom_point(data=data_rect_hf_ne$BATCH1,mapping=aes(y1,y2,color=ID,shape=ID),size=0.1)+
  geom_path(data=data_rect_hf_ne$BATCH1,mapping=aes(y1,y2,color=ID,shape=ID),size=0.1)+
  geom_point(data = data_rect_hf_ne$BATCH1%>% filter(!duplicated(ID)),
             aes(x = y1, y = y2), shape = 3, size = 4, col = "red")+
  xlab("x") + ylab("y")

ggsave(filename="samples_rect_hf_ne.png",plot=plot_samples_rect_hf_ne,width=10,height=5)
ggplotly(plot_samples_rect_hf_ne)
```

@@

#### Violin plot of the estimated coefficients for one batch

@@

for (k in 1:N_BATCH) {
  m<-crcvm_rect_hf_ne1[[paste("BATCH",k,sep="")]]
  coeffs=rbind(m$coeff_re(),m$coeff_fe())
  coeff_names=rownames(coeffs)
  coeff_values=as.numeric(coeffs)
  sub_index=grepl("omega",coeff_names)
  sub_coeffs_df=data.frame("estimate"=coeff_values[sub_index],"coeff_name"=factor(coeff_names[sub_index]))
  coeffs_df=rbind(coeffs_df,sub_coeffs_df)
}

@@

### Biais and RMSE

```{r compute biais and rmse}
biais=c()
rmse=c()
for (string in names(sp_coeff_ExpShore)) {
  name=paste("omega.",string,sep="")
  true_value=sp_coeff_ExpShore[string]
  estimates=coeffs_df[coeffs_df$coeff_name==name,"estimate"]
  biais=c(biais,1/N_BATCH*sum(abs(estimates-true_value))/true_value)
  rmse=c(rmse,sqrt(1/N_BATCH*sum((estimates-true_value)^2)))
}

names(biais)=paste("omega.",names(sp_coeff_ExpShore),sep="")
names(rmse)=paste("omega.",names(sp_coeff_ExpShore),sep="")
```

```{r print biais}
biais
```

```{r print rmse}
rmse
```

@@

```{r plot estimates te DistanceShore for one batch of low error samples}
xmin=list("theta"=-pi,"DistanceShore"=0.3)
xmax=list("theta"=pi,"DistanceShore"=2)
xlabel=list("DistanceShore"="Distance to shore","theta"="Theta")
link=list("DistanceShore"=identity,"theta"=identity)
true_smooths=list("omega"=fomega_splines)
res=crcvm_rect_hf_ne2$BATCH1$get_all_plots(baseline=NULL,model_name="crcvm_rect_hf_ne2",xmin=xmin,xmax=xmax,
                                    xlabel=xlabel,link=link,show_CI="pointwise",true_smooths=true_smooths,save=TRUE)
res$fe_omega_theta_DistanceShore
```

@@

  #fitted model
  m<-crcvm_rect_hf_ne2[[paste("BATCH",k,sep="")]]

  #re and fe coeffs
  coeffs=rbind(m$coeff_re(),m$coeff_fe())

  #names and values of coeffs
  coeff_names=rownames(coeffs)
  coeff_values=as.numeric(coeffs)

  #keep coeffs that do not contain ID
  sub_index=grep("ID",coeff_names,invert=TRUE)

  #standard deviation of random effects
  sdev=m$sdev()
  sdev_names=rownames(sdev)[1:2]
  sdev_values=as.numeric(sdev)[1:2]

  #bind dataframe
  sub_coeffs_df=data.frame("estimate"=c(coeff_values[sub_index],sdev_values),"coeff_name"=factor(c(coeff_names[sub_index],sdev_names)))
  coeffs_df=rbind(coeffs_df,sub_coeffs_df)
}

@@

p=ggplotly(p)
saveWidget(p, file ="crcvm_rect_hf_ne2/all_batches_coeffs_crcvm_rect_hf_ne1.html")

### Biais and RMSE

```{r compute biais and rmse DistanceShore splines}
biais=c()
rmse=c()

#biais and rmse for omega coefficients
for (string in names(sp_coeff_Dshore)) {
  name=paste("omega.",string,sep="")
  true_value=sp_coeff_Dshore[string]
  estimates=coeffs_df[coeffs_df$coeff_name==name,"estimate"]
  biais=c(biais,1/N_BATCH*sum(abs(estimates-true_value)))
  rmse=c(rmse,sqrt(1/N_BATCH*sum((estimates-true_value)^2)))
}

#add biais and rmse for tau
true_log_tau=0
estimates=coeffs_df[coeffs_df$coeff_name=="tau.(Intercept)","estimate"]
biais=c(biais,1/N_BATCH*sum(abs(estimates-true_log_tau)))
rmse=c(rmse,sqrt(1/N_BATCH*sum((estimates-true_log_tau)^2)))

#add biais and rmse for nu
true_log_nu=log(4)
estimates=coeffs_df[coeffs_df$coeff_name=="nu.(Intercept)","estimate"]
biais=c(biais,1/N_BATCH*sum(abs(estimates-true_log_nu)))
rmse=c(rmse,sqrt(1/N_BATCH*sum((estimates-true_log_nu)^2)))

#add biais and rmse for sdev tau
true_sdev_tau=0.1
estimates=coeffs_df[coeffs_df$coeff_name=="tau.(ID)","estimate"]
biais=c(biais,1/N_BATCH*sum(abs(estimates-true_sdev_tau)))
rmse=c(rmse,sqrt(1/N_BATCH*sum((estimates-true_sdev_tau)^2)))

#add biais and rmse for sdev tau
true_sdev_nu=0.3
estimates=coeffs_df[coeffs_df$coeff_name=="nu.(ID)","estimate"]
biais=c(biais,1/N_BATCH*sum(abs(estimates-true_sdev_nu)))
rmse=c(rmse,sqrt(1/N_BATCH*sum((estimates-true_sdev_nu)^2)))

names(biais)=c(paste("omega.",names(sp_coeff_Dshore),sep=""),"tau.(Intercept)","nu.(Intercept)","tau.(ID)","nu.(ID)")
names(rmse)=c(paste("omega.",names(sp_coeff_Dshore),sep=""),"tau.(Intercept)","nu.(Intercept)","tau.(ID)","nu.(ID)")

```{r print biais DistanceShore splines}
biais
```

```{r print rmse DistanceShore splines}
rmse
```

@@

fDshore=function(z,v,p) {

@@

  #distance to shore
  Dshore=sqrt(normal[1]^2+normal[2]^2)
  return (Dshore)
}

@@

atw=list("DistanceShore"=fDshore,"AngleNormal"=fangle)
sim_data=crcvm_rect_hf_ne2$BATCH1$simulate(z0=x0[1,],data=crcvm_rect_hf_ne2$BATCH1$data(),atw=atw,land=domain)
